The \project{unRAVE} catalog

### 2016-08-24 article/unrave.tex
Finish description of the training sets

\epic\ differs from the \project{Kepler} input catalog because \epic\ does not 
benefit from having narrow-band $DDO_{51}$ photometry in order to aid dwarf/giant 
classification.  Despite this limitation the labels in the \epic\ catalog have 
already been shown to be accurate and trustworthy \citep{Huber_2016}.  However, 
when the posteriors are wide (i.e., the quoted confidence intervals are large) 
due to limited information available, it is possible that a star has been 
mis-classified.  This is most prevalent for sub-giants, where \citet{Huber_2016} 
note that $\approx55-70$\% of sub-giants are misclassified as dwarfs.  The 
probability of misclassification is usually quantified in the uncertainties of 
each star; most sub-giants that could be dwarfs have large confidence intervals.  
Therefore, requiring low uncertainties will decrease the total sample size, but 
will remove most misclassifications.  The situation is far more favourable for 
dwarfs and giants.  Only 1-4\% of giant stars are misclassified as dwarfs, and 
about 7\% of dwarfs are misclassified as giants.  To summarise, the \epic\ labels 
with narrow confidence intervals are usually of high fidelity, and any spurious 
misclassifications would be clearly distinguishable if spectra were available.

We sought to have a small overlap between our giant and main-sequence star training
sets.  Most of our giant training set is encapsulated within $0 < \logg < 3.5$, 
however there is a sparse sampling of stars reaching to $\logg \approx 4$.  For
the \epic\ main-sequence/sub-giant star training set we required $\logg > 3.5$,
allowing for $\approx0.5$~dex of overlap between the two training sets.  We further
employed the following quality constraints: the upper and lower confidence intervals 
in $\teff$ must be below 150~K; the upper and lower confidence intervals in $\logg$ 
must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must exceed 
30~per~pixel$^{-1}$; and $\teff \leqslant 6750$.  Unfortunately these strict constraints
removed most metal-poor stars, which we later found to cause the test labels to have
under-predicted abundances for dwarfs of low metallicity.  For this reason we relaxed
(ignored) those quality constraints for stars in the \epic\ catalog with 
$[{\rm Fe/H}] < -1$.  Of the stars we forcibly included with $[{\rm Fe/H}] < -1$,
none were classified as being sub-giants in \epic, therefore they have a low (7\%)
likelihood of misclassification.  The distilled sample of the \rave--\epic\ overlap
catalog contains 583 stars (of 4,611).

% We can also identify misclassified stars by cross-validation.

% Make sure you mention the lack of abundance information in K2/EPIC

% Including Giant stars of low metallicity.
% Refer to the training set figure.
% Worry later about label offsets.

>>>>

\epic\ differs from the \project{Kepler} input catalog because \epic\ does not 
benefit from having narrow-band $DDO_{51}$ photometry in order to aid dwarf/giant 
classification.  Despite this limitation the labels in the \epic\ catalog have 
already been shown to be accurate and trustworthy \citep{Huber_2016}.  However, 
when the posteriors are wide (i.e., the quoted confidence intervals are large) 
due to limited information available, it is possible that a star has been 
mis-classified.  This is most prevalent for sub-giants, where \citet{Huber_2016} 
note that $\approx55-70$\% of sub-giants are misclassified as dwarfs.  The 
probability of misclassification is usually quantified in the uncertainties of 
each star; most sub-giants that could be dwarfs have large confidence intervals.  
Therefore, requiring low uncertainties will decrease the total sample size, but 
will remove most misclassifications.  The situation is far more favourable for 
dwarfs and giants.  Only $1-4$\% of giant stars are misclassified as dwarfs, and 
about 7\% of dwarfs are misclassified as giants.  To summarise, the \epic\ labels 
with narrow confidence intervals are usually of high fidelity, and any spurious 
misclassifications would be clearly distinguishable if spectra were available.

We sought to have a small overlap between our giant and main-sequence star training
sets.  Most of our giant training set is encapsulated within $0 < \logg < 3.5$, 
however there is a sparse sampling of stars reaching to $\logg \approx 4$.  For
the \epic\ main-sequence/sub-giant star training set we required $\logg > 3.5$,
allowing for $\approx0.5$~dex of overlap between the two training sets.  We further
employed the following quality constraints: the upper and lower confidence intervals 
in $\teff$ must be below 150~K; the upper and lower confidence intervals in $\logg$ 
must be less than 0.15~dex; the $S/N$ of the \rave\ spectra must exceed 
30~per~pixel$^{-1}$; and $\teff \leqslant 6750$~K.  Unfortunately these strict constraints
removed most metal-poor stars, which we later found to cause the test labels to have
under-predicted abundances for dwarfs of low metallicity.  For this reason we relaxed
(ignored) those quality constraints for stars in the \epic\ catalog with 
$[{\rm Fe/H}] < -1$.  Of the stars we forcibly included with $[{\rm Fe/H}] < -1$,
none were classified as being sub-giants in \epic, therefore they have a low (7\%)
likelihood of misclassification.  After training a model based on main-sequence and
giant stars (Section \ref{sec:the-model}), we found we could further identify 
misclassifications by leave-one-out cross-validation (iteratively, even).  However,
we chose not to do this because the number of likely misclassifications in the training
set was negligible ($\approx1$\%), and the improvement in main-sequence test set labels
was minimal.  The distilled sample of the \rave--\epic\ overlap catalog contains 583 
stars (of 4,611).

We previously alluded to including giant stars from \epic\ in our training set based 
on \apogee\ labels.  This was principally to ensure an adequate balance of metal-poor
giant stars in the training set.  Without this inclusion, we again found that the
abundances of giant stars would be under-predicted at low metallicity, critically 
affecting any inferences we wish to make with low metallicity stars (e.g., globular
clusters).  The metallicities in the \epic\ catalog come from a variety of spectroscopic
sources, or photometrically-selected when no spectroscopic information is available.
While these heterogeneous metallicities may introduce scaling differences, on balance we
later found that it was more important to include reasonably accurate (albeit imprecise)
metallicities of metal-poor stars in the training set rather than excluding metal-poor
stars entirely.  In Section~\ref{sec:validation} we show that these potential issues
have a negligible impact on the test set labels, and we further show that our abundances
are consistent with high-resolution studies in the literature.

We have constructed a justified training set for stars across the main-sequence, sub-giant,
and red giant branch.  However the lack of overlap between \rave\ and other works have
resulted in a somewhat peculiar situation. Detailed abundances are available from \apogee\
for all giant stars in our sample, but only imprecise metallicities are available from
\epic\ for stars on the main-sequence and the sub-giant branch.  In the following section
we will construct a simplified model that does not include detailed abundances, before
demonstrating how we (1) improve upon the imprecise metallicities for main-sequence/sub-giant
stars, and (2) translate abundances from the giant branch stars to the main-sequence/sub-giant
branch sample.
